---
title: "Macquarie_Island_case_study"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

This tutorial is derived from the document *Tutorial-2-Macquarie_Island_case_study* written in Python environment by Kristensen *et.al.* (2019) (https://github.com/nadiahpk/qualitative-modelling/tree/master/tutorials). Using the same data, this document aims to give a better illustration of the similarities and the slight differences between the original qualitative modeling code written in Python and this R version. 

```{r include=FALSE}
library(LogicOpt)
source('qualmod.R')
source('findPCU.R')
```

## 1. Define the species interaction network

Macquarie Island interaction network si defined by the species present, the interactions between them, and the signs of these interactions.

```{r}
sppList = c(
    'albatrosses',
    'prions',
    'burrowSeabirds',
    'petrels',
    'herbfield',
    'macroInverts',
    'mice',
    'penguins',
    'rabbits',
    'rats',
    'redpolls',
    'skuas',
    'surfaceSeabirds',
    'tussock'
)

positive_edges_list = list(
    'prions'= c('grassland'),
    'skuas'= c('prions','burrowSeabirds','rabbits','penguins'),
    'petrels'= c('penguins','tussock','grassland'),
    'mice'= c('herbfield','macroInverts','tussock'),
    'rats'= c('macroInverts','herbfield','tussock'),
    'burrowSeabirds'= c('tussock'),
    'rabbits'= c('tussock','herbfield','grassland'),
    'macroInverts'= c('herbfield','grassland','tussock'),
    'albatrosses'= c('tussock','herbfield'),
    'redpolls'= c('macroInverts','tussock','herbfield','grassland')
)

negative_edges_list = list(
    'prions'= c('prions','skuas'),
    'skuas'= c('skuas','tussock'),
    'penguins'= c('penguins','skuas','petrels'),
    'petrels'= c('petrels'),
    'mice'= c('mice','rats'),
    'rats'= c('rats'),
    'burrowSeabirds'= c('burrowSeabirds','skuas','rabbits'),
    'rabbits'= c('rabbits','skuas'),
    'surfaceSeabirds'= c('surfaceSeabirds','rats'),
    'macroInverts'= c('macroInverts','rats','mice','redpolls'),
    'tussock'= c('tussock','mice','rats','rabbits'),
    'albatrosses'= c('albatrosses'),
    'herbfield'= c('herbfield','rabbits'),
    'grassland'= c('grassland'),
    'redpolls'= c('redpolls')
)
```

The function initialise_foodweb returns a DiagrammeR graph obect, storing information of the network structure (nodes and edges) as dataframes (i.e. NDF: node data frame, and EDF: edge data frame). The network can be quickly plotted using function *render_graph()* from package *DiagrammeR* for a quick check of network structures. For a neat plot, use funciton *foodweb_neat_plot()*.  

```{r initialse_network_macquarie}
web <- initialise_foodweb(positive_edges_list, negative_edges_list)
render_graph(web, title = "foodweb_all_links")
```

```{r neat_plot_macquarie}
foodweb_neat_plot(web, title = "mq island foodweb neat")
```

Using the function *qualitative_community_matrix*, we can convert the interaction network into a qualitative community matrix (Mq), and get two named vectors *labelToIndex* and *indexToLabel* to map species labels (i.e. names) to indices of matrix and vice versa.  

```{r}
outputQM <- qualitative_community_matrix(web)
Mq <- outputQM$Mq
Mq
```

```{r}
labelToIndex <- outputQM$labelToIndex
indexToLabel <- outputQM$indexToLabel

unname(indexToLabel[10])
unname(labelToIndex["rabbits"])
```

We encode the validation criteria and then use the function *get_conditions_df* to get a dataframe with three columns: species' names (label), their corresponding conditions and their corresponding indices in the qualitative community matrix. We also get the indices of the controled species and the indices of a list of species (*sppList*) that we want to monitor.

```{r}
# response to increase in rabbits
validation <- list(
    "rabbits" = 1,
    "tussock"= -1)

condn_df <- get_condition_df(labelToIndex,validation)
condn_df

control_list <- c("rabbits")
control_list_idx <- unname(labelToIndex[control_list])

sppList_idx <- unname(labelToIndex[sppList])
```

## 2. Probabilistic approach: perform Monte Carlo Simulation.

We run 10^3^ times of simulations.

*(The original tutorial does not perform the Monte Carlo simulation to this specific case study.)*

```{r}
set.seed(178)

noSim <- 1000
collectedResponses = list()

sz <- dim(Mq)
n <- length(Mq)

start_time <- Sys.time()

for (i in 1:noSim){
  
    valid <- FALSE
    
    while (!valid) {
      
      # find a random community matrix that is stable
      maxEig = 1
      
      while (maxEig > 0) {
        
          M = matrix(runif(n), sz[1], sz[2]) * Mq
          maxEig <- max(Re(eigen(M, symmetric=FALSE, only.values=TRUE)$values))
      }
      
      # Now have a valid stable matrix
      # find the sensitivity matrix
      Sq <- -solve(M)
      
      # check validation criteria
      valid <- all(sign(Sq[condn_df$idx, unname(labelToIndex["rabbits"])]) == condn_df$conditions)
    }
    
    #Now have a valid stable community matrix
    response <- vector()
    
    for (ps in control_list_idx) {
      
        resp <- ifelse(Sq[sppList_idx, ps] < 0, "neg", 
                       ifelse(Sq[sppList_idx, ps] >0, "pos", "zer"))
        
        response <- append(response, resp)
    }
    
    collectedResponses[[i]] <- response
  
}

end_time <- Sys.time()
time_elapsed = end_time - start_time
print(time_elapsed)
```

Convert the collected list of species responses into a dataframe.  
```{r}
df_responses <- do.call(rbind, collectedResponses) %>% as.data.frame() %>% mutate_if(is.factor, as.character)

# short labels
colnames <- unlist(lapply(control_list, function(x) paste0(str_sub(x, 1, 3), "_", str_sub(sppList, 1, 3))))

# full labels
# colnames <- unlist(lapply(control_list, function(x) paste0(x, "_", sppList)))

colnames(df_responses) <- colnames
head(df_responses)
```

Aggregate simulation outcomes.  

```{r}
levels <- c('pos', 'neg') 
df_responses[] <-  lapply(df_responses, factor, levels=levels)

count <- sapply(df_responses, table)
count
```

## 3.The Boolean approach

### 3.1 Parameter sweep

To save time in this tutorial, we hard-coded the number of species-response combinations to be found, which is 36. If the number is not known beforehand, you can specify the number of matrices to generate using the commented-out *for* loop code.

```{r}
collectedResponses = list()
sz <- dim(Mq)
n <- length(Mq)

start_time <- Sys.time()

while (length(collectedResponses) < 36) {

# noSim <- 1000000 
# for (i in 1:noSim){
    
    valid <- FALSE
    
    # find a random community matrix that is stable
    while (!valid) {
      
        # find a random community matrix that is stable
        maxEig = 1
        
        while (maxEig > 0) {
          
          M = matrix(runif(n), sz[1], sz[2]) * Mq
          maxEig <- max(Re(eigen(M, symmetric=FALSE, only.values=TRUE)$values))
        }
        
        # Now have a valid stable matrix
        # find the sensitivity matrix
        Sq <- -solve(M)
        
        # check validation criteria
        valid <- all(sign(Sq[condn_df$idx, unname(labelToIndex["rabbits"])]) == condn_df$conditions)
    }
    
    #Now have a valid stable community matrix
    response <- vector()
    
    for (ps in control_list_idx) {
      
        resp <- ifelse(Sq[sppList_idx, ps] < 0, "neg", 
                       ifelse(Sq[sppList_idx, ps] >0, "pos", "zer"))
        
        response <- append(response, resp)
    }
    
    if (!any(collectedResponses %in% list(response))) {
      
        collectedResponses[[length(collectedResponses)+1]] <- response
    }
}

end_time <- Sys.time()
time_elapsed = end_time - start_time
print(time_elapsed)

length(collectedResponses)
```

Write the parameter-sweep results (i.e. unobserved species-response combinations) into a csv file.  

```{r}
df_responses <- do.call(rbind, collectedResponses)
colnames <- unlist(lapply(control_list, function(x) paste0(x, "_", sppList))) # full labels
colnames(df_responses) <- colnames

fileNames <- unlist(lapply(control_list, function(x) paste0("uniques_web1_", x, ".csv")))
write.csv(df_responses, file = fileNames, row.names=FALSE)
```

### 3.2 Boolean analysis

Read in the responses from the csv file that was written previously.
```{r}
df_mq <- read.csv("uniques_web1_rabbits.csv", head = TRUE)
# head(df_mq)

str4true = 'pos'
str4flase = 'neg'

allResponse <- colnames(df_mq); 
allResponse
```

We write a list of desired responses, *desiredResponses*. The *desiredResponses* list is then converted into a Boolean mask *desiredResponsesMask*, and passed to the function *getUnobservedInts2*.  

Function *getUnobservedInts* finds the complement of the set of observed responses (i.e. unobserved response combinations), returning a list of unobserved responses as a list of integers.

```{r}
desiredResponses = c(
    'rabbits_albatrosses',
    'rabbits_prions',
    'rabbits_burrowSeabirds',
    'rabbits_petrels',
    'rabbits_herbfield',
    'rabbits_macroInverts',
    'rabbits_mice',
    'rabbits_penguins',
    'rabbits_rats',
    'rabbits_redpolls',
    'rabbits_skuas',
    'rabbits_surfaceSeabirds')

boolLen = length(desiredResponses)

desiredResponsesMask <- which(allResponse %in% desiredResponses)

unobservedInts <- getUnobservedInts(df_mq, desiredResponsesMask, boolLen, str4true)

length(unobservedInts)
```

The function *getUnobservedBooldf* turns the list of integers *unobservedInts*, corresponding to unobserved species responses, back into a dataframe in Boolean expression (1s and 0s). The function *getUnobservedBooldf* also add an new column *unob* which is a vector of 1s (the dataframe therefore becomes a truth table) to allow funciton *logicopt()* in Package *LogicOpt* to perform Boolean minimization on the dataframe.  

```{r}
unobservedBooldf <- getUnobservedBooldf(unobservedInts, desiredResponses)
# unobservedBooldf[50:55, 7:13] # check
```

The function*logicopt()* in Package *LogicOpt* is used to perform the Boolean minimisation.  

```{r}
start_time <- Sys.time()

opt <- logicopt(unobservedBooldf, boolLen, 1, mode="espresso")
optEqn <- tt2eqn(opt[[1]], boolLen, 1)
# optEqn   # show the ON set equations for the minimized truth table

end_time <- Sys.time()
time_elapsed = end_time - start_time
print(time_elapsed)
```

The function *getPCUList* converts the optimized equations (*optEqn_c*) into a list of strings (PCUs) for further analysis.  

```{r}
PCUList <- getPCUList(optEqn, str4true, str4flase, desiredResponses)
PCUList
```


```{r}
niceNames = rbind(
        c('rabbits', 'rabbits'),
        c('petrels', 'petrels'),
        c('mice', 'mice'),
        c('burrowSeabirds', 'burrow-nest seabirds'),
        c('macroInverts', 'macroinvertebrates'),
        c('herbfield', 'herbfield'),
        c('redpolls', 'redpolls'),
        c('skuas', 'skuas'),
        c('rats', 'rats'),
        c('surfaceSeabirds', 'surface-nest seabirds'),
        c('penguins', 'penguins'),
        c('prions', 'prions'),
        c('albatrosses', 'albatrosses'),
        c('tussock', 'tussock')
)

niceNames <- setNames(niceNames[ ,2], niceNames[ ,1]) #create a named vector as a dictionary
```


PCULists are used as inputs to the functions *get_edgelist_singleAnte* , which get an edgelist in a single-antecedent form. Then the edgelist will be passed to *draw_implication_network* function to plot the corresponding implication network.

```{r implication_network_macquarie}
edgelist <- get_edgelist_singleAnte(PCUList)
# edgelist
draw_implication_network(edgelist, niceNames)
```




















