---
title: "Christmas_Island_complicated_implication_networks"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
date: "21 June 2019"
---

Boolean analysis for the combined cat and rat management for the full Town Network.

```{r include=FALSE}
source('findPCU.R')
library(LogicOpt)
```

Read in the csv file of observed species-response combinations that we saved in advance for the full run of the parameter-sweep (number of matrices produced: 10^8^).

```{r read_species_response_csv}
observedResponse_df <- read.csv("town_search_full_10e8.csv", head = TRUE)
nrow(observedResponse_df)
```

# 1. Perform Boolean minimisation and get the PCUList

Performing the Boolean minimization on species responses to the combined management. Write a list of desired responses for this management.  

```{r Boolean_minimization_combined_management}
allResponse <- colnames(observedResponse_df)
# allResponse # check species response

str4true = 'pos'
str4flase = 'neg'

desiredResponses_cr = c(
  'cat_tropicBird', 
  'cat_goshawk', 
  'cat_feralChicken', 
  'cat_hawkOwl', 
  'cat_rat', 
  'cat_flyingFox',
  'rat_tropicBird', 
  'rat_goshawk', 
  'rat_feralChicken', 
  'rat_hawkOwl', 
  'rat_cat', 
  'rat_flyingFox')

boolLen_cr <- length(desiredResponses_cr)
desiredResponsesMask_cr <- match(desiredResponses_cr, allResponse)

unobservedInts_cr <- getUnobservedInts2(observedResponse_df, desiredResponsesMask_cr, boolLen_cr)

unobservedBooldf_cr <- getUnobservedBooldf(unobservedInts_cr, desiredResponses_cr)

opt_cr <- logicopt(unobservedBooldf_cr, boolLen_cr, 1, mode="espresso")
optEqn_cr <- tt2eqn(opt_cr[[1]], boolLen_cr, 1)

PCUList_cr <- getPCUList(optEqn_cr, str4true, str4flase, desiredResponses_cr)
```

Since the PCUList is quite large and complicated, we split the PCUList up by length.  

```{r split_PCU_by_length}
PCUlength_cr <- sapply(PCUList_cr, function(i) length(i))
PCUList_cr1 <- PCUList_cr[PCUlength_cr <= 5]
PCUList_cr2 <- PCUList_cr[PCUlength_cr > 5]
```

Draw an implication network for all rules with length less than 5.  

```{r network_all_rules_less_than_five}
edgelist_cr1 <- get_edgelist_certainAnte(PCUList_cr1, c())
draw_implication_network(edgelist_cr1)
```

Above implication network shows that 'poscat_rat' 'negcat_rat' and 'posrat_cat' are central contingencies. Thus we analyze the implication rules under different scenarios based on the rat response to cat management and cat response to rat management.  

# 2. Implication netwoks under different scenarios

First, we split up the PCUs by the effects (positive/negtaive) of cat management on rat population. The always true conditions (here are "when cat management has a positive effect on rat population" and "when cat management has a negative effect on rat population" for PCULList_cr3 and PCUList_cr4 respectively) will be shown as "True" for a succinct presentation. There are rules only involving species-responses to cat management, which have been predicated by the cat only management scenario (see the document *Christmas_Island_Boolean_approach*). Thus, we can perform a further pruning, and these rules are predicated on "True".  

```{r split_PCUList_poscat_rat_and_negcat_rat}
PCUList_cr3 = list() 
PCUList_cr4 = list()
  
for (PCU in PCUList_cr1){  
    if (!('negcat_rat' %in% PCU)){
        if(!all(str_detect(PCU, "cat"))) {
          
            PCU <- str_subset(PCU, 'cat_rat', negate = TRUE)
            PCUList_cr3[[length(PCUList_cr3)+1]] <- PCU
        }
    } 
}

for (PCU in PCUList_cr1){  
    if (!('poscat_rat' %in% PCU)){
        if(!all(str_detect(PCU, "cat_"))) {
          
            PCU <- str_subset(PCU, 'cat_rat', negate = TRUE)
            PCUList_cr4[[length(PCUList_cr4)+1]] <- PCU
        }
    } 
}

```


## Scenario 1: when cat management has a positive effect on rat population.  

```{r scenario1_poscat_rat_all_rules}
edgelist_cr3 <- get_edgelist_certainAnte(PCUList_cr3, c())
draw_implication_network(edgelist_cr3)
```

From above implication network, we could further explore the implication rules under two scenarios given the effect of rat managemnet on cat population.


### Scenario 1a: when cat management has a positive effect on rat population and rat management has a negative effect on cat population (this is the most likely scenario).  

```{r scenario1a_poscat_rat_and_negrat_cat_PCUList}
PCUList_cr5 = list() 
  
for (PCU in PCUList_cr3){  
    if (!('posrat_cat' %in% PCU)){
        PCU <- str_subset(PCU, 'rat_cat', negate = TRUE)
        PCUList_cr5[[length(PCUList_cr5)+1]] <- PCU
    }
}
```


The always true conditions (here are "when cat management has a positive effect on rat population and rat management has a negative effect on cat population") are shown as "True" for a succinct presentation.  

In addtion, there is a better way to present the network given that the effect of rat management upon the hawk-owl was contingent upon the effect of cat management upon the hawk-owl.

```{r scenario1a_poscat_rat_and_negrat_cat}
edgelist_cr5 <- get_edgelist_certainAnte(PCUList_cr5, c('poscat_hawkOwl', 'negcat_hawkOwl'))
draw_implication_network(edgelist_cr5)
```

### Scenario 1b: when cat management has a positive effect on rat population and rat management has a positive effect on cat population (this is a less likely scenario).  

```{r scenario1b_poscat_rat_and_posrat_cat_PCUList}
PCUList_cr6 = list()
for (PCU in PCUList_cr3){  
    if (!('negrat_cat' %in% PCU)){
        PCU <- str_subset(PCU, 'rat_cat', negate = TRUE)
        PCUList_cr6[[length(PCUList_cr6)+1]] <- PCU
    }
}
```

```{r scenario1b_poscat_rat_and_posrat_cat}
edgelist_cr6 <- get_edgelist_certainAnte(PCUList_cr6, c('poscat_hawkOwl', 'negcat_hawkOwl'))
draw_implication_network(edgelist_cr6)
```

## Scenario 2: when cat management has a negative effect on rat population.  

```{r scenario2_negcat_rat_all_rules}
edgelist_cr4 <- get_edgelist_certainAnte(PCUList_cr4, c())
draw_implication_network(edgelist_cr4)
```

### Scenario 2a: When cat management has a negative effect on rat population and rat management has a negative effect on cat population.  

```{r scenario2a_negcat_rat_negrat_cat_PCUList}
PCUList_cr7 = list()
for (PCU in PCUList_cr4){  
    if (!('posrat_cat' %in% PCU)){
        PCU <- str_subset(PCU, 'rat_cat', negate = TRUE)
        PCUList_cr7[[length(PCUList_cr7)+1]] <- PCU
    }
}
```

```{r scenario2a_negcat_rat_negrat_cat}
edgelist_cr7a <- get_edgelist_certainAnte(PCUList_cr7, c('poscat_hawkOwl', 'negcat_hawkOwl'))
draw_implication_network(edgelist_cr7a)
```

### Scenario 2b: when cat management has a negative effect on rat population and rat management has a positive effect on cat population.  

```{r scenario2b_negcat_rat_posrat_cat_PCUList}
PCUList_cr8 = list()
for (PCU in PCUList_cr4){  
    if (!('negrat_cat' %in% PCU)){
        PCU <- str_subset(PCU, 'rat_cat', negate = TRUE)
        PCUList_cr8[[length(PCUList_cr8)+1]] <- PCU
    }
}
```

```{r scenario2b_negcat_rat_posrat_cat}
edgelist_cr8a <- get_edgelist_certainAnte(PCUList_cr8, c('poscat_hawkOwl', 'negcat_hawkOwl'))
draw_implication_network(edgelist_cr8a)
```
