---
title: "Christmas_Island_probabilistic_approach"
date: 21 June 2019
output:
  pdf_document: default
  html_notebook: default
---

In this document, we define the species interaction network for the Town region on Chirstmas Island. Then we perform the probabilistic modeling approach and species' response to different management perturbations are pooled. 

```{r include=FALSE}
source('qualmod.R')
```

## 1. Define the species interaction network
We define the Christmas Island Town network by the species present, the interactions between them, and the signs of these interactions. The function initialise_foodweb returns a DiagrammeR graph obect, storing information of the network structure (nodes and edges) as dataframes (i.e. NDF: node data frame, and EDF: edge data frame). The network can be quickly plotted using function *render_graph()* from package *DiagrammeR* for a quick check of network structures. For a neat plot, use funciton *foodweb_neat_plot()*.  
  
  
```{r define_interaction_network}
spp_list = c(
  'cat',
  'rat',
  'crab',
  'goshawk',
  'hawkOwl',
  'tropicBird',
  'flyingFox',
  'feralChicken',
  'kestrel',
  'groundCultivars',
  'canopyCultivars',
  'diurnalInsectResources',
  'nocturnalInsectResources'
)

positive_edges_list = list(
    'cat'= c('tropicBird', 'flyingFox', 'diurnalInsectResources', 'feralChicken', 'rat'),
    'rat'= c('tropicBird', 'groundCultivars', 'canopyCultivars', 'diurnalInsectResources', 'nocturnalInsectResources', 'feralChicken'),
    'crab'= c('groundCultivars'),
    'goshawk'= c('rat', 'tropicBird', 'diurnalInsectResources', 'feralChicken'),
    'hawkOwl'= c('rat', 'nocturnalInsectResources'),
    'flyingFox'= c('canopyCultivars'),
    'feralChicken'= c('diurnalInsectResources'),
    'kestrel'= c('rat', 'diurnalInsectResources')
)

negative_edges_list = list(
    'cat'= c('cat'),
    'rat'= c('rat', 'cat', 'crab', 'goshawk', 'hawkOwl', 'kestrel'),
    'crab'= c('crab'),
    'goshawk'= c('goshawk'),
    'hawkOwl'= c('hawkOwl'),
    'tropicBird'= c('tropicBird', 'cat', 'rat', 'goshawk'),
    'flyingFox'= c('flyingFox', 'cat'),
    'feralChicken'= c('feralChicken', 'cat', 'rat', 'goshawk'),
    'kestrel'= c('kestrel'),
    'groundCultivars'= c('groundCultivars', 'crab', 'rat'),
    'canopyCultivars'= c('canopyCultivars', 'flyingFox', 'rat'),
    'diurnalInsectResources'= c('diurnalInsectResources', 'cat', 'rat', 'goshawk', 'feralChicken', 'kestrel'),
    'nocturnalInsectResources'= c('nocturnalInsectResources', 'rat', 'hawkOwl')
)

unmonitored_spp_list = c(
    'crab', 
    'kestrel', 
    'groundCultivars', 
    'canopyCultivars', 
    'diurnalInsectResources', 
    'nocturnalInsectResources')

unmonitored_spp_list_sim = c(
    'cat',
    'rat',
    'feralChicken',
    'crab', 
    'kestrel', 
    'groundCultivars', 
    'canopyCultivars', 
    'diurnalInsectResources', 
    'nocturnalInsectResources')

control_list = c('cat','rat')

ambig_edges_list = list(
    list('rat'= c('crab')),
    list(
        'rat'= c('kestrel'),
        'kestrel'= c('rat')
    ) 
    # Remove interactions between rat and kestrel, which indicates the ambiguous links 
    # between these two species were keeped or removed together. However, there could 
    # be cases that the bidirectional link were considered as two ambiguous links respectively, 
    # In such case, the two links will not be wrapped together, and thus they could be treated 
    # respectively. 
)

```

Initialise the full Town network.  
  
  
```{r initialise_network}
full_web <- initialise_foodweb(positive_edges_list, negative_edges_list) 
render_graph(full_web) # for a quick plot
```

Get a neat plot of the full Town network.  
  
  
```{r network_neat_plot}
foodweb_neat_plot(full_web, title = "CI Town full web")
```

Using the function *qualitative_community_matrix*, we can convert the interaction network into a qualitative community matrix (Mq), and get two named vectors *labelToIndex* and *indexToLabel* to map species labels (i.e. names) to indices of matrix and vice versa.  
  
  
```{r qualitative_community_matrix}
output <- qualitative_community_matrix(full_web)
Mq <- output$Mq
Mq
```

```{r}
labelToIndex <- output$labelToIndex
indexToLabel <- output$indexToLabel

unname(indexToLabel[10])
unname(labelToIndex["kestrel"])
```
  
Get following vectors:
- *spp_list_idx* represents indices of species in *spp_list* in order..
- *control_list_idx* represents indices of pest species in *control_list* in order.
- *monitored_spp_list_sim* is a vector of species being monitored in the probabilistic approach.
- *monitored_spp_idx_sim* represents indices of species in *monitored_spp_list_sim*.  
  
  
```{r}
spp_list_idx <- unname(labelToIndex[spp_list])
control_list_idx <- unname(labelToIndex[control_list])

monitored_spp_list_sim <- spp_list[!spp_list %in% unmonitored_spp_list_sim]
monitored_spp_idx_sim <- unname(labelToIndex[monitored_spp_list_sim])
```

## 2.The probabilistic approach: randomly sampling perturbation responses 

The following *for* loop run simulations to collect species responses. To save time in this illustration, we run 10^5^ times of simulations.  
  
Ambiguous links are removed randomly from the network, which results in different network structures. Modelling outcomes of different network structures are pooled. All the simulation outcomes are stored in a list *collectedResponses*.  
  
  
```{r Monte_Carlo_simulation}
set.seed(178)

nSim <- 10000

collectedResponses = list()
sz <- dim(Mq)
n <- length(Mq)
noAmbig <- length(ambig_edges_list)
    
start_time <- Sys.time()

for (i in 1: nSim){
    
    selectAmbig <- as.logical(rbinom(noAmbig, 1, runif(1))) # randomly select ambig.links to remove

    if (all(!selectAmbig)) {
      
        Mq = Mq
        
    } else {
      
        ambig_df<- 
            ambig_edges_list[selectAmbig] %>%
            melt() %>%
            setNames(., c("labelfrom", "labelto", "list")) %>%
            mutate_if(is.factor, as.character)
        
        dropEdge <- cbind(unname(labelToIndex[ambig_df[, "labelto"]]), unname(labelToIndex[ambig_df[, "labelfrom"]]))
        Mq[dropEdge] <- rep(0, nrow(ambig_df)) # set those dropped links as 0s.
    }
  
    
    valid <- FALSE
    
    while (!valid) {
      
      # find a random community matrix that is stable
      maxEig = 1
      
      while (maxEig > 0) {
        
          M = matrix(runif(n), sz[1], sz[2]) * Mq
          maxEig <- max(Re(eigen(M, symmetric=FALSE, only.values=TRUE)$values))
      }
      
      # Now have a valid stable matrix, find the sensitivity matrix
      Sq <- -solve(M)
      
      # check validation criteria: the pest respond negatively to management
      control_easy_cat = (Sq[labelToIndex['cat'],labelToIndex['cat']] > 0) 
      control_easy_rat = (Sq[labelToIndex['rat'],labelToIndex['rat']] > 0)
      valid <- all(control_easy_cat, control_easy_rat)
    }
    
    #Now have a valid stable community matrix
    response <- vector()
    
    for (ps in control_list_idx) {
      
        resp <- ifelse(-Sq[monitored_spp_idx_sim, ps] < 0, "neg", "pos")
        
        response <- append(response, resp)
    }
    
    collectedResponses[[i]] <- response
  
}

end_time <- Sys.time()
time_elapsed = end_time - start_time
print(time_elapsed)
```

Convert the list *collectedResponses* storing simulation outcomes into a dataframe *df_responses*.  
  
  
```{r dataframe_of_species_responses}
df_responses <- do.call(rbind, collectedResponses) %>% as.data.frame() %>% mutate_if(is.factor, as.character)

colnames <- unlist(lapply(control_list, function(x) paste0(x, "_", monitored_spp_list_sim)))
colnames(df_responses) <- colnames
head(df_responses,3) # check the dataframe.
```

## 3.Aggregte outcomes for different managements

Aggregate outcomes for cat management only.  
  
  
```{r outcomes_of_cat_management}
df_cat_resp <- df_responses %>%
    select(., starts_with("cat"))
count_cat <- sapply(df_cat_resp, table)
count_cat
```

Aggregate outcomes for the combined cat and rat management.  
  
  
```{r outcomes_of_combined_management}
combined_resp_list <- list()

for (sp in monitored_spp_list_sim){
  
    # For each species, combine its response to cat and response to rat in each of the simulations.
    combined_response <- 
        cbind(select(df_responses, contains(sp))) %>%
        apply(., 1 , paste , collapse = "&" )
    
    combined_resp_list[[length(combined_resp_list)+1]] <- combined_response
}

df_combined_resp <- 
    do.call(cbind, combined_resp_list) %>% 
    as.data.frame()
  
colnames_combined_df <- unlist(lapply(monitored_spp_list_sim, function(x) paste0('cr_', x)))
colnames(df_combined_resp) <- colnames_combined_df

lvls_cr <- c('pos&pos', 'pos&neg', 'neg&pos','neg&neg') 
df_combined_resp[] <-  lapply(df_combined_resp, factor, levels=lvls_cr)

count_cr <- sapply(df_combined_resp, table)
count_cr

```
